Trajectory Mapping Results

### 2020-12-30 scripts/analyze_trajectories_v2.Rmd
wip dataset trajectories

```{r debugging}
# Debugging
args$burnin = 0.1
args$demes = read.csv("/Users/maceci/code/mt-analysis/201014_europe2/files/demes.csv", stringsAsFactors = FALSE)
args$input = "/Users/maceci/code/mt-analysis/europeGLM/results/europeC/201106_europe5.TL.traj"
args$metadata =  "/Users/maceci/code/mt-analysis/europeGLM/results/europeC/final_metadata.tsv"
args$n = 200
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
demes$deme[3] = "China"
demes$division[3] = NA
demes <- demes %>% arrange(deme)
```

@@

zoom_to <- c(20, 50) 
zoom_level <- 2.6
target_crs <- sprintf('+proj=aeqd +lon_0=%f +lat_0=%f',
                      zoom_to[1], zoom_to[2])
C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^zoom_level
y_span <- C / 2^(zoom_level+1)

@@

events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = ifelse(is.na(dest), as.character(src),
                       as.character(factor(dest,
                                           levels = 0:(nrow(demes) - 1),
                                           labels = sort(demes$deme)))),
         date_model = date(date_decimal(decimal_date(ymd(MRS)) - age))) %>%
  mutate(date = date_model + 10)
```

@@

# Summary statistics
gt_summary <- gt %>%
  group_by(date, type) %>%
  summarize(Imean = mean(N),
            Imedian = median(N),
            Ilow    = quantile(N, 0.025), #HPD??
            Ihigh   = quantile(N, 0.975),
            .groups = "drop_last") %>%
  ungroup()

# Grid events, time grid of 1 day, taking the sum of events
ge <- gridEventsByAge(events, ages_1day) %>%
  mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
         date = date_model + 10,
         dest = ifelse(is.na(dest), as.character(src), as.character(dest)))

@@

>>>>

```{r debugging}
# Debugging
args$burnin = 0.1
args$demes = read.csv("./files/demes.csv", stringsAsFactors = FALSE)
args$input = "./results/europeC/201106_europe5.TL.traj"
args$metadata =  "./results/europeC/final_metadata.tsv"
args$n = 50
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
demes$deme[3] = "China"
demes$division[3] = NA
demes <- demes %>% arrange(deme)
```

@@

zoom_to <- c(20, 50) 
zoom_level <- 2.5
target_crs <- sprintf('+proj=aeqd +lon_0=%f +lat_0=%f',
                      zoom_to[1], zoom_to[2])
C <- 40075016.686   # ~ circumference of Earth in meters
x_span <- C / 2^zoom_level
y_span <- C / 2^(zoom_level+1)

@@

events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = ifelse(is.na(dest), as.character(src),
                       as.character(factor(dest,
                                           levels = 0:(nrow(demes) - 1),
                                           #labels = sort(demes$deme),
                                           labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain")))),
         date_model = date(date_decimal(decimal_date(ymd(MRS)) - age))) %>%
  mutate(date = date_model + 10,
         src = ifelse(event == "O", "China", src),
         dest = ) #TODO
```

@@

# Summary statistics
gt_summary <- gt %>%
  group_by(date, type) %>%
  summarize(Imean = mean(N),
            Imedian = median(N),
            Ilow    = quantile(N, 0.025), #95% CI, compute HPD??
            Ihigh   = quantile(N, 0.975),
            .groups = "drop_last") %>%
  ungroup()

# Grid events, time grid of 1 day, taking the sum of events
ge <- gridEventsByAge(events, ages_1day) %>%
  mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
         date = date_model + 10,
         dest = ifelse(is.na(dest), as.character(src), as.character(dest))) %>%
  ungroup %>%
  filter(N != 0)

@@

```{r}
# Active cases
gt %>%
    # 1. Prepare data
    filter(type == deme) %>%
    mutate(traj = as.character(traj)) %>%
    bind_rows(case_data %>%
                rename(type = deme,
                       N = cumcases) %>%
                filter(type == deme) %>%
                mutate(traj = "ecdc") %>%
                select(type, traj, date, N) %>%
                distinct()) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, N, group = traj,
                  color = traj == "ecdc",
                  alpha = traj == "ecdc",
                  linetype = traj == "ecdc",
                  size = traj == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(deme, " inferred\n population trajectories"),
                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                         " ECDC total case count")),
                       values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 2),
                          labels = c(paste0(deme, " inferred\n population trajectories"),
                                     paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                            " ECDC total case count"))) +
    scale_size_manual(name = "", values = c(0.5, 1),
                      labels = c(paste0(deme, " inferred\n population trajectories"),
                                 paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                        " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())

```{r}
# Cumulative cases
events <- df$events

events[events$event == "O", "src"] = 2
events[events$event == "O", "mult"] =  1

df_cases1 <- events %>%
  group_by_at(vars(-mult)) %>%
  summarise(N = sum(mult), .groups = "drop") %>%
  mutate(event2 = ifelse(event == "M", paste0("IM_", src), NA),
         event3 = ifelse(event == "M", paste0("OM_", dest), NA)) %>%
  pivot_longer(src:dest, names_to = "role", values_to = "type") %>%
  mutate(event = ifelse(event == "M" & role == "src", event3,
                        ifelse(event == "M" & role == "dest", event2, event))) %>% 
  select(-c("role", "event2", "event3")) %>%
  filter(!is.na(type)) %>%
  group_by_at(vars(-N)) %>%
  summarise(value = sum(N), .groups = "drop") %>%
  pivot_wider(names_from = event, values_from = value, values_fill = list(value = 0)) %>%
  mutate(IM = IM_0 + IM_1 + IM_2 + IM_3 + IM_4 + IM_5,
         OM = OM_0 + OM_1 + OM_2 + OM_3 + OM_4 + OM_5) %>%
  group_by(traj, type) %>%
  arrange(traj, type, time) %>%
  mutate_at(tail(names(.), -4), list(c = cumsum)) %>%
  mutate(activePop = O_c + B_c + IM_c - D_c - OM_c - S_c,
         cumPop = O_c + B_c + IM_c,
         inPop = O + B + IM,
         outPop = D + OM + S) %>% 
  pivot_longer(O:outPop, names_to = "var", values_to = "value")
  # pivot_longer(O:outPop, names_to = c("value", "set"),
  #  names_pattern = "(.*)_(.)")

df_cases <- events %>%
  group_by_at(vars(-mult)) %>%
  summarise(N = sum(mult), .groups = "drop") %>%
  mutate(srctmp = src, destmp = dest) %>%
  pivot_longer(src:dest, names_to = "role", values_to = "type") %>%
  mutate(event = ifelse(event == "M" & role == "src", "OM",
                        ifelse(event == "M" & role == "dest", "IM", event)),
         partner = ifelse(event == "OM", destmp,
                        ifelse(event == "IM", srctmp, NA))) %>% 
  select(-c("role", "srctmp", "destmp")) %>%
  filter(!is.na(type)) %>%
  group_by_at(vars(-N)) %>%
  summarise(value = sum(N), .groups = "drop") %>%
  group_by(traj, type, partner, event) %>%
  arrange(traj, time, type) %>%
  mutate(cumvalue = cumsum(value))

mig_cum = df_cases %>%
  filter(event %in% c("IM", "OM")) %>%
  group_by(traj, event, time, age, type) %>%
  summarise(value1 = sum(value), .groups = "drop") %>%
  group_by(traj, event, type) %>%
  arrange(traj, time, type) %>%
  mutate(cumvalue1 = cumsum(value1),
            partner1 = "all")

inout_pop = df_cases %>% 
  mutate(var = ifelse(event %in% c("O", "B", "IM"), "in_pop", "out_pop")) %>%
  group_by(traj, type, var, time, age) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  group_by(traj, type, var) %>%
  arrange(traj, time, type) %>%
  mutate(cumvalue = cumsum(value))

active_pop = inout_pop %>%
  pivot_wider(names_from = var, values_from = value, values_fill = list(value = 0)) %>%
  mutate(active_pop = in_pop - out_pop) %>%
  pivot_longer(active_pop, names_to = "var", values_to = "value") %>%
  group_by(traj, type) %>%
  arrange(traj, time, type) %>%
  mutate(cumvalue = cumsum(value)) %>%
  select(-c("in_pop", "out_pop"))

df_cases <- df_cases %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = ifelse(is.na(dest), as.character(src),
                       as.character(factor(dest,
                                           levels = 0:(nrow(demes) - 1),
                                           #labels = sort(demes$deme),
                                           labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain")))))

ge %>%
    # 1. Prepare data
    filter((dest == deme & event == "B") | (dest == deme & event == "M")) %>%
  filter(traj == 1) %>%
    mutate(traj = as.character(traj)) %>%
    bind_rows(case_data %>%
                rename(dest = deme,
                       N = cumcases) %>%
                filter(dest == deme) %>%
                mutate(traj = "ecdc") %>%
                select(dest, traj, date, N) %>%
                distinct()) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, N, group = traj,
                  color = traj == "ecdc",
                  alpha = traj == "ecdc",
                  linetype = traj == "ecdc",
                  size = traj == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(deme, " inferred\n population trajectories"),
                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                         " ECDC total case count")),
                       values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 2),
                          labels = c(paste0(deme, " inferred\n population trajectories"),
                                     paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                            " ECDC total case count"))) +
    scale_size_manual(name = "", values = c(0.5, 1),
                      labels = c(paste0(deme, " inferred\n population trajectories"),
                                 paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
                                        " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())
